The Radio Sky at Meter Wavelengths: $m$-Mode Analysis Imaging with the Owens Valley Long Wavelength Array

### 2017-06-29 paper.tex
fixup figures
now using matplotlib 2 so the defaults aren't so horrendous anymore
add an example l curve

@@

The process of Tikhonov regularization minimizes $\|\b v - \b B\b a\|^2 + \epsilon\|\b a\|^2$ for
some arbitrary value of $\epsilon > 0$ chosen by the observer. The solution that minimizes this
expression is given by
\begin{equation}\label{eq:tikhonov-solution}
    \atikh = (\b B^*\b B + \epsilon\b I)^{-1}\b B^*\b v\,.
\end{equation}
Tikhonov regularization adds a small value $\epsilon$ to the diagonal of $\b B^*\b B$, fixing the
matrix's singularity. We can see this by using the singular value decomposition (SVD) of the matrix
$\b B = \b U \b \Sigma \b V^*$. Equation~\ref{eq:tikhonov-solution} becomes
\begin{equation}
    \atikh = \b V (\b\Sigma^2 + \epsilon \b I)^{-1}\b\Sigma \b U^*\b v\,,
\end{equation}
where
\[
    \b\Sigma = \left(
        \begin{array}{ccc}
            \sigma_1 & & \\
                     & \sigma_2 & \\
                     & & \ddots \\
        \end{array}
    \right)\,.
\]
The diagonal elements of $\b\Sigma$ are the singular values of $\b B$ and so the contribution of
each singular component to the Tikhonov-regularized solution is scaled by $\sigma_i / (\sigma_i^2 +
\epsilon)$, where $\sigma_i$ is the singular value for the $i$th singular component. Tikhonov
regularization therefore acts to suppress any component for which $\sigma_i\lesssim\sqrt{\epsilon}$.
If $\sigma_i = 0$, the component is set to zero.

In practice the measurement $\b v$ is corrupted by noise with covariance $\b N$. For illustrative
purposes we will assume that $\b N=n\b I$ for some scalar $n$. In this case the covariance of the
Tikhonov-regularized spherical harmonic coefficients is
\begin{equation}
    \b C = n \b V (\b\Sigma^2 + \epsilon\b I)^{-2} \b\Sigma^2 \b V^*\,.
\end{equation}
Each singular component is scaled by a factor of $\sigma_i^2/(\sigma_i^2 + \epsilon)^2$.
In the absence of Tikhonov regularization ($\epsilon=0$)
singular components with the smallest singular values -- the ones that the interferometer is
the least sensitive to -- actually come to dominate the covariance of the measured spherical
harmonic coefficients. Tikhonov regularization improves this situation by downweighting these
components.

While Tikhonov regularization will force unmeasured modes to zero, if a prior map of the sky already
exists, it will be preferrable to instead minimize $\|\b v - \b B\b a\|^2 + \epsilon \|\b a-\b
a_\text{prior}\|^2$.

@@

Tikhonov regularization requires the observer to pick the value of $\epsilon$. If $\epsilon$ is too
large then too much importance is placed on minimizing the norm of the solution and the
least-squares residuals will suffer. However if $\epsilon$ is too small then the problem will be
poorly regularized and the resulting sky map may not represent the true sky. Picking the value of
$\epsilon$ therefore requires understanding the trade-off between the two norms.

This trade-off can be analyzed quantitatively by trialing several values of $\epsilon$ and computing
$\|\b v - \b B\b a\|^2$ and $\|\b a\|^2$ for each trial. An example is shown in Figure \todo{create
a figure}. The shape of this curve has a characteristic L-shape, and as a result this type of plot
is called an L-curve \todo{reference}. The ideal value of $\epsilon$ lies near the turning point of
the plot. At this point a small decrease in $\epsilon$ will lead to an undesired rapid increase in
$\|\b a\|^2$, and a small increase in $\epsilon$ will lead to an undesired rapid increase in $\|\b v
- \b B\b a\|^2$.

In practice, the L-curve should be used as a guide to estimate a reasonable value of $\epsilon$.
However better results can often be obtained by tuning the value of $\epsilon$. For instance
increasing the value of $\epsilon$ can improve the noise properties of the map by down-weighting
noisy modes. Decreasing the value of $\epsilon$ can improve the resolution in the map by
up-weighting the contribution of longer baselines, which are likely fewer in number. In this respect
choosing the value of $\epsilon$ is analagous to picking the weighting scheme in traditional
imaging where robust weighting schemes can be tuned to similar effect \citep{briggs}.

@@

While $m$-mode analysis imaging automatically deconvolves large angular scales, point sources do not
get deconvolved by the imaging process. This is a limitation of the set of basis functions we have
chosen to represent the sky -- the spherical harmonics. Point sources carry power to large values of
$l$, but the regularization will eventually begin to downweight spherical harmonics with $l$ larger
than what the interferometer is sensitive to. This consequence of this is that point sources will
appear in the maps with some characteristic point spread function. This point spread function may be
computed with
\begin{equation}
    \b a_\text{PSF}(\theta, \phi)
        = (\b B^*\b B + \epsilon\b I)^{-1}\b B^*\b B\b a_\text{PS}(\theta, \phi)\,,
\end{equation}
where $\b a_\text{PSF}(\theta, \phi)$ is the vector of spherical harmonic coefficients representing
the point spread function at the spherical coordinates $(\theta, \phi)$, and $\b a_\text{PS}(\theta,
\phi)$ is the vector of spherical harmonic coefficients for a point source at $(\theta, \phi)$ given
by
\begin{equation}
    a_{lm, \text{PS}}(\theta, \phi) = Y_{lm}^*(\theta, \phi)\,.
\end{equation}

@@

%\begin{figure*}[ht]
%    \includegraphics[width=\textwidth]{maps/rgb-map-small}
%    \caption{three color map}
%\end{figure*}
%
%\begin{figure*}[ht]
%    \centering
%    \begin{tabular}{cc}
%        \includegraphics[width=\columnwidth, draft]{maps/rgb-map} &
%        \includegraphics[width=\columnwidth, draft]{maps/rgb-map} \\
%        \includegraphics[width=\columnwidth, draft]{maps/rgb-map} &
%        \includegraphics[width=\columnwidth, draft]{maps/rgb-map} \\
%        \includegraphics[width=\columnwidth, draft]{maps/rgb-map} &
%        \includegraphics[width=\columnwidth, draft]{maps/rgb-map} \\
%        \includegraphics[width=\columnwidth, draft]{maps/rgb-map} &
%        \includegraphics[width=\columnwidth, draft]{maps/rgb-map} \\
%    \end{tabular}
%    \caption{single channel maps}
%\end{figure*}
%
%\begin{figure*}[ht]
%    \centering
%    \begin{tabular}{cc}
%        \includegraphics[width=\columnwidth, draft]{maps/rgb-map} &
%        \includegraphics[width=\columnwidth, draft]{maps/rgb-map} \\
%        \includegraphics[width=\columnwidth, draft]{maps/rgb-map} &
%        \includegraphics[width=\columnwidth, draft]{maps/rgb-map} \\
%        \includegraphics[width=\columnwidth, draft]{maps/rgb-map} &
%        \includegraphics[width=\columnwidth, draft]{maps/rgb-map} \\
%        \includegraphics[width=\columnwidth, draft]{maps/rgb-map} &
%        \includegraphics[width=\columnwidth, draft]{maps/rgb-map} \\
%    \end{tabular}
%    \caption{jackknifed error maps}
%\end{figure*}

>>>>

\begin{figure*}[ht]
    \includegraphics[width=\textwidth]{maps/rgb-map-small}
    \caption{three color map}
    \label{fig:rgb}
\end{figure*}

@@

The process of Tikhonov regularization minimizes $\|\b v - \b B\b a\|^2 + \varepsilon\|\b a\|^2$ for
some arbitrary value of $\varepsilon > 0$ chosen by the observer. The solution that minimizes this
expression is given by
\begin{equation}\label{eq:tikhonov-solution}
    \atikh = (\b B^*\b B + \varepsilon\b I)^{-1}\b B^*\b v\,.
\end{equation}
Tikhonov regularization adds a small value $\varepsilon$ to the diagonal of $\b B^*\b B$, fixing the
matrix's singularity. We can see this by using the singular value decomposition (SVD) of the matrix
$\b B = \b U \b \Sigma \b V^*$. Equation~\ref{eq:tikhonov-solution} becomes
\begin{equation}
    \atikh = \b V (\b\Sigma^2 + \varepsilon \b I)^{-1}\b\Sigma \b U^*\b v\,,
\end{equation}
where
\[
    \b\Sigma = \left(
        \begin{array}{ccc}
            \sigma_1 & & \\
                     & \sigma_2 & \\
                     & & \ddots \\
        \end{array}
    \right)\,.
\]
The diagonal elements of $\b\Sigma$ are the singular values of $\b B$ and so the contribution of
each singular component to the Tikhonov-regularized solution is scaled by $\sigma_i / (\sigma_i^2 +
\varepsilon)$, where $\sigma_i$ is the singular value for the $i$th singular component. Tikhonov
regularization therefore acts to suppress any component for which
$\sigma_i\lesssim\sqrt{\varepsilon}$.  If $\sigma_i = 0$, the component is set to zero.

In practice the measurement $\b v$ is corrupted by noise with covariance $\b N$. For illustrative
purposes we will assume that $\b N=n\b I$ for some scalar $n$. In this case the covariance of the
Tikhonov-regularized spherical harmonic coefficients is
\begin{equation}
    \b C = n \b V (\b\Sigma^2 + \varepsilon\b I)^{-2} \b\Sigma^2 \b V^*\,.
\end{equation}
Each singular component is scaled by a factor of $\sigma_i^2/(\sigma_i^2 + \varepsilon)^2$.  In the
absence of Tikhonov regularization ($\varepsilon=0$) singular components with the smallest singular
values -- the ones that the interferometer is the least sensitive to -- actually come to dominate
the covariance of the measured spherical harmonic coefficients. Tikhonov regularization improves
this situation by downweighting these components.

While Tikhonov regularization will force unmeasured modes to zero, if a prior map of the sky already
exists, it will be preferrable to instead minimize $\|\b v - \b B\b a\|^2 + \varepsilon \|\b a-\b
a_\text{prior}\|^2$.

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=\columnwidth]{figures/lcurve/lcurve}
    \caption{
        Example l-curve (from spw04)
    }
    \label{fig:lcurve}
\end{figure}

Tikhonov regularization requires the observer to pick the value of $\varepsilon$. If $\varepsilon$
is too large then too much importance is placed on minimizing the norm of the solution and the
least-squares residuals will suffer. However if $\varepsilon$ is too small then the problem will be
poorly regularized and the resulting sky map may not represent the true sky. Picking the value of
$\varepsilon$ therefore requires understanding the trade-off between the two norms.

This trade-off can be analyzed quantitatively by trialing several values of $\varepsilon$ and
computing $\|\b v - \b B\b a\|^2$ and $\|\b a\|^2$ for each trial. An example is shown in Figure
\todo{create a figure}. The shape of this curve has a characteristic L-shape, and as a result this
type of plot is called an L-curve \todo{reference}. The ideal value of $\varepsilon$ lies near the
turning point of the plot. At this point a small decrease in $\varepsilon$ will lead to an undesired
rapid increase in $\|\b a\|^2$, and a small increase in $\varepsilon$ will lead to an undesired
rapid increase in $\|\b v - \b B\b a\|^2$.

In practice, the L-curve should be used as a guide to estimate a reasonable value of $\varepsilon$.
However better results can often be obtained by tuning the value of $\varepsilon$. For instance
increasing the value of $\varepsilon$ can improve the noise properties of the map by down-weighting
noisy modes. Decreasing the value of $\varepsilon$ can improve the resolution in the map by
up-weighting the contribution of longer baselines, which are likely fewer in number. In this respect
choosing the value of $\varepsilon$ is analagous to picking the weighting scheme in traditional
imaging where robust weighting schemes can be tuned to similar effect \citep{briggs}.

@@

While $m$-mode analysis imaging automatically deconvolves large angular scales, point sources do not
get deconvolved by the imaging process. This is a limitation of the set of basis functions we have
chosen to represent the sky -- the spherical harmonics. Point sources carry power to large values of
$l$, but the regularization will eventually begin to downweight spherical harmonics with $l$ larger
than what the interferometer is sensitive to. This consequence of this is that point sources will
appear in the maps with some characteristic point spread function. This point spread function may be
computed with
\begin{equation}
    \b a_\text{PSF}(\theta, \phi)
        = (\b B^*\b B + \varepsilon\b I)^{-1}\b B^*\b B\b a_\text{PS}(\theta, \phi)\,,
\end{equation}
where $\b a_\text{PSF}(\theta, \phi)$ is the vector of spherical harmonic coefficients representing
the point spread function at the spherical coordinates $(\theta, \phi)$, and $\b a_\text{PS}(\theta,
\phi)$ is the vector of spherical harmonic coefficients for a point source at $(\theta, \phi)$ given
by
\begin{equation}
    a_{lm, \text{PS}}(\theta, \phi) = Y_{lm}^*(\theta, \phi)\,.
\end{equation}

@@

%\begin{figure*}[ht]
%    \centering
%    \begin{tabular}{cc}
%        \includegraphics[width=\columnwidth, draft]{maps/rgb-map} &
%        \includegraphics[width=\columnwidth, draft]{maps/rgb-map} \\
%        \includegraphics[width=\columnwidth, draft]{maps/rgb-map} &
%        \includegraphics[width=\columnwidth, draft]{maps/rgb-map} \\
%        \includegraphics[width=\columnwidth, draft]{maps/rgb-map} &
%        \includegraphics[width=\columnwidth, draft]{maps/rgb-map} \\
%        \includegraphics[width=\columnwidth, draft]{maps/rgb-map} &
%        \includegraphics[width=\columnwidth, draft]{maps/rgb-map} \\
%    \end{tabular}
%    \caption{single channel maps}
%\end{figure*}
%
%\begin{figure*}[ht]
%    \centering
%    \begin{tabular}{cc}
%        \includegraphics[width=\columnwidth, draft]{maps/rgb-map} &
%        \includegraphics[width=\columnwidth, draft]{maps/rgb-map} \\
%        \includegraphics[width=\columnwidth, draft]{maps/rgb-map} &
%        \includegraphics[width=\columnwidth, draft]{maps/rgb-map} \\
%        \includegraphics[width=\columnwidth, draft]{maps/rgb-map} &
%        \includegraphics[width=\columnwidth, draft]{maps/rgb-map} \\
%        \includegraphics[width=\columnwidth, draft]{maps/rgb-map} &
%        \includegraphics[width=\columnwidth, draft]{maps/rgb-map} \\
%    \end{tabular}
%    \caption{jackknifed error maps}
%\end{figure*}
